\documentclass{ximera}

\input{../preamble.tex}

\title{Waves on Transmission Lines}

\begin{document}

\begin{abstract}
%Stuff can go here later if we want!
\end{abstract}

\maketitle

\begin{sectionOutcomes}

After completing this section, students should be able to do the following.

\begin{itemize}
\item Name several transmission lines.
\item Explain the difference between phase shift and time delay of a sinusoidal signal.
\item Evaluate whether the transmission line theory or circuit theory has to be used based on the length of the line and the frequency
\item Students will explain the difference between lumped and distributed circuit elements.
\item Derive the voltage on a transmission line from a consideration of a time-delay due to the finite speed of signals in a transmission-line circuit.
\item Explain parts of propagation constant and what they represent.
\item Calculate the phase and attenuation constant for specific transmission lines.
\item Identify whether the wave travels in the positive or negative direction from the equation of a wave.
\item Describe how signal flows on a transmission line
\item Recognize and explain transmission line equivalent circuit model
\item Derive the equations for voltage and current waves on a transmission line from the equivalent circuit model.
\item Describe forward and reflected wave on a transmission line.
\item Sketch forward and reflected wave as a function of distance, and explain how the graph changes as time passes.
\item Derive phasor form of voltage and current wave from/to the time-domain form.
\item Describe what wavelength represents on a graph of a wave vs. distance
\item Explain how the wavelength is similar and different to wave’s period?
\item Derive and calculate the transmission line impedance and reflection coefficient.
\item Relate reflection coefficient to impedance.
\item Derive and calculate the input impedance of a transmission line
\item Calculate and visualize phasors of forward going voltage and current waves at various points on a transmission line.
\end{itemize}

\end{sectionOutcomes}

\end{document}
